---
title: 'Reproducible Research: Peer Assessment 1'
author: "Samuel Bozzi Baco"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

# 0 LOADING PACKAGES

To run the analysis, the package *tidyverse* will be loaded.

```{r load_package, results = 'hide',message= FALSE}
library(tidyverse)
library(knitr)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


# 01 LOADING AND PRE PROCESSING DATA

To read the zip file directly, the package *readr* will be used. The function was configured using the Rstudio UI option, and the code was replicated here so the reproducibility is maintained.

```{r load_data}
activity <- read_csv("activity.zip", col_types = cols(steps = col_double()))
```

## Summary the data

All variables for activity were summarized together to save space. There are three variables on the dataframe: (1) **steps** (number of steps taking in a 5-minute interval, (2) **date** (the date on which the measurement was taken in YYYY-MM-DD format) and (3) **interval** (identifier for the 5-minute interval in which measurement was take).

```{r summary}
summary(activity)
```

# 02 MEAN STEPS TAKEN BY DAY

Splitting the data by different days, the average value of steps is 9354 and the median is 10395 steps.

```{r steps_total}
stepDay <- activity %>%
  group_by(date) %>%
  summarise(total = sum(steps, na.rm = TRUE))

summary(stepDay$total)
```
```{r steps_daily_hist}

ggplot(stepDay, aes(total)) +
  geom_histogram(fill = "dodgerblue4") +
  geom_vline(aes(xintercept = mean(total), col = "darkred"), size = 1, linetype = 4 ) +
  geom_vline(aes(xintercept = median(total), col = "green"), size = 1, linetype = 4 ) +
  scale_color_identity(guide = "legend",
                       name = "Stats from Steps",
                       labels = c("Mean", "Median")) +
  labs(x = "Steps", 
       y = "",
       title = "Counts of Steps",
       subtitle = "Daily Basis") +
  theme_minimal() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "antiquewhite4"),
        axis.title = element_text(colour = "antiquewhite4"),
        axis.text = element_text(colour = "antiquewhite4"),
        plot.title = element_text(colour = "antiquewhite4"),
        plot.subtitle = element_text(colour = "antiquewhite4"),
        strip.text = element_text(colour = "antiquewhite4"))
```

# 03 AVERAGE STEPS FOR 5 MINUTES INTERVAL 

Below there is a time series of the averages of steps taken for all 5 minutes interval across all days. The maximum average of steps happens at 835 minutes, with a value of 206 (rounded). 

```{r steps_intervals}

stepInterval <- activity %>%
  group_by(interval) %>%
  summarise(mean = mean(steps, na.rm = TRUE))

ggplot(stepInterval, aes(x = interval, y = mean)) +
  geom_line(colour = "dodgerblue4", size = 1) +
  labs(x = "5 minute interval", 
       y = "",
       title = "Average of Steps") +
  theme_minimal() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "antiquewhite4"),
        axis.title = element_text(colour = "antiquewhite4"),
        axis.text = element_text(colour = "antiquewhite4"),
        plot.title = element_text(colour = "antiquewhite4"),
        plot.subtitle = element_text(colour = "antiquewhite4"),
        strip.text = element_text(colour = "antiquewhite4")) +
  geom_point(data = stepInterval[which.max(stepInterval$mean), ], color = "darkred", size = 3) +
  geom_text(data = stepInterval[which.max(stepInterval$mean), ], color = "darkred", size = 4,
            label = paste(round(stepInterval$mean[which.max(stepInterval$mean)]), "steps at", stepInterval$interval[which.max(stepInterval$mean)], "min", sep = " "),
            hjust = -0.1)

```


# 04 TREATING MISSING VALUES

The NA’s values from variable **steps** are distributed along whole dataset and represent 13% of all data (2304 data points).

````{r na_summary}
stepNA <- is.na(activity$steps)

summary(activity$date[stepNA])
```
```{r na_sum}
sum(stepNA)
```

```{r na_mean}
mean(stepNA)
```
```{r na_plot_date}

activity %>%
  filter(is.na(steps)) %>%
  ggplot(aes(date)) +
  geom_histogram(fill = "dodgerblue4") +
  labs(x = "Date", 
       y = "",
       title = "Counts of NA's for Step Variable") +
  theme_minimal() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "antiquewhite4"),
        axis.title = element_text(colour = "antiquewhite4"),
        axis.text = element_text(colour = "antiquewhite4"),
        plot.title = element_text(colour = "antiquewhite4"),
        plot.subtitle = element_text(colour = "antiquewhite4"),
        strip.text = element_text(colour = "antiquewhite4"))

```

The missing values were replaced by the median of the steps, calculated by each 5 minute interval. 

```{r imputed_summary}
stepMedian <- activity %>%
  group_by(interval) %>%
  summarise(median = median(steps, na.rm = TRUE))

stepImp <- activity %>%
  left_join(stepMedian) 

stepImp$steps[is.na(stepImp$steps)] <- stepImp$median[is.na(stepImp$steps)]

stepDay2 <- stepImp %>%
  group_by(date) %>%
  summarise(total = sum(steps, na.rm = TRUE))

summary(stepDay2$total)
```

```{r imputed_plo}

ggplot(stepDay2, aes(total)) +
  geom_histogram(fill = "dodgerblue4") +
  geom_vline(aes(xintercept = mean(total), col = "darkred"), size = 1, linetype = 1 ) +
  geom_vline(aes(xintercept = median(total), col = "green"), size = 1, linetype = 1 ) +
  geom_vline(data = stepDay, aes(xintercept = mean(total), col = "darkred"), size = 1, linetype = 4) + 
  geom_vline(data = stepDay, aes(xintercept = median(total), col = "green"), size = 1, linetype = 4) + 
  scale_color_identity(guide = "legend",
                       name = "Stats",
                       labels = c("Mean", "Median")) +
  labs(x = "Steps", 
       y = "",
       title = "Counts of Steps",
       subtitle = "Daily Basis - Dashed Line are not imputaded") +
  theme_minimal() +
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "antiquewhite4"),
        axis.title = element_text(colour = "antiquewhite4"),
        axis.text = element_text(colour = "antiquewhite4"),
        plot.title = element_text(colour = "antiquewhite4"),
        plot.subtitle = element_text(colour = "antiquewhite4"),
        strip.text = element_text(colour = "antiquewhite4"))

```

# 05 ACTIVITIE DIFFERENCE BETWEEN WEEK DAYS AND WEEK ENDS

The *weekdays()* function was used to determine the weekdays name for **date** variable. The result was stored at **wday** variable. Then, an *if.else* flow was used to transform the character value from **wday** to a two level factor, stored in **wknd**. The imputed data frame was used to this computations. 

After the factor creation, the steps were summarized by mean, grouped by 5 minute interval and week day or week end. 

```{r weekend}
stepImp$wday <- weekdays(stepImp$date)

stepImp$wknd <- as.factor(ifelse(stepImp$wday == "sábado" |
                                   stepImp$wday == "domingo", "Weekend", "Weekday"))

stepWeekend <- stepImp %>%
  group_by(interval, wknd) %>%
  summarise(mean = mean(steps, na.rm = TRUE))

stepMedian <- stepWeekend %>%
  group_by(wknd) %>%
  summarise(median = median(mean))

stepMean <- stepWeekend %>%
  group_by(wknd) %>%
  summarise(mean = mean(mean))
```

```{r weekend plot}
ggplot(stepWeekend, aes(x = interval, y = mean)) + 
  facet_grid(. ~ wknd) +
  geom_line(color =  "dodgerblue4", size = 1) +
  geom_hline(data = stepMedian, aes(yintercept = median, color = "orange"), size = 1.5) +
  geom_hline(data = stepMean, aes(yintercept = mean, color = "grey"), size = 1.5) +
  scale_color_identity(guide = "legend",
                       name = "Steps stats",
                       labels = c("Total Mean", "Total Median")) +
  labs(title = "Comparison of activity level between weekdays and weekends",
       subtitle = "Mean of steps",
       y = "",
       x = "5 minute interval") +
  theme_minimal() + 
  theme(panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "antiquewhite4"),
        axis.title = element_text(colour = "antiquewhite4"),
        axis.text = element_text(colour = "antiquewhite4"),
        plot.title = element_text(colour = "antiquewhite4"),
        plot.subtitle = element_text(colour = "antiquewhite4"),
        strip.text = element_text(colour = "antiquewhite4"),
        legend.text = element_text(colour = "antiquewhite4"),
        legend.title = element_text(colour = "antiquewhite4")
  )
```

The general acrivity seems to be a liitle higher on weekend, no matter if you consider the total mean of total median.


