---
title: "README"
author: "Samuel Bozzi Baco"
date: "29/02/2020"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, results = 'hide', include = FALSE}

library(dplyr)

setwd("D:/Samuel/Meus Documentos/Google Drive/R/Learning/Johns_Hopkins_Coursera/Course_03_Getting_Cleaning_Data/Project")

load("run_analysis_environment.RData")

```


## General Explanation

This document explains the analysis for Samsung Accelerometer data, asked to complete the activity for final project of Coursera’s Johns Hopkins Getting and Cleaning data.

The code run_analysis has seven (7) sections, being the section “0” related to opening files. In this section both *subjects* and *activities* variables receive names, to not be confounded whit other variables during the analysis. Also, specific objects for variables names (*acLabels* and *featNames*) are created.


```{r}
names(yTrain) <- "activity"
```
```{r}
names(subjTrain) <- "subject"
```
## Section 1: Merging Datasets

The section 1 has the code to merges both test and train data sets. The results are stored at *dfComplete* dataframe.

```{r}
dfComplete <- rbind(train, test)
```
## Section 2: Extracting only Mean and Standard Deviation variables

The section 2 look for patterns related to Mean and Standard deviation at variable names (stored at vector *featNames*) and select only those, storing the result at *dfSelected* dataframe.

```{r}
meanVar <- grep("mean\\(\\)", featNames$V2)

sdVar <- grep("std\\(\\)", featNames$V2)

dfSelected <- dfComplete[ ,c(meanVar, sdVar, 562:563)]
```

## Section 3: Renaming activities

The section 3 renames all values at *activites* variable using the function *mapvalues* from *plyr* package. 

```{r}
dfSelected$activity <- plyr::mapvalues(dfSelected$activity , 
                                   from = sort(unique(dfSelected$activity)),
                                   to = acLabels$V2)
```
## Section 4: Column names

The section 4 renames all the variables on *dfSelected* dataframe with original names found at databook supplied. 

```{r tidy = FALSE}
colnames(dfSelected)[colnames(dfSelected) %in% 
                             c(names(dfSelected[ ,-c(67:68)]))] <- featNames$V2[c(meanVar, sdVar)]
```

## Section 5: Average by subjects and activities

The section 5 calculates the averages by subjects and activities, using dyplr *pipes* and functions. First, the dataframe is grouped by subjects and activities, than all averages are calculated.

```{r}
average <- dfSelected %>%
        group_by(subject, activity) %>%
        summarise_all(mean)
```

## Section 6: Write dataframe

The section 6 has the code to write the dataframe as a *.txt* file.

```{r}
write.table(average, file = "average.txt", row.name = FALSE)
```









