---
title: "Notebook for Statistical Inference Course Project 02"
author: "Samuel B Baco"
output:
  pdf_document: default
  html_notebook: default
---
```{r, message=F, warning=F}
library(tidyverse)
```


# PART 02: BASIC INFERENTIAL DATA ANALYSIS

This part consist on analysing the ToothGrowth dataset from *datasets* package. Below there is a summary of the dataset.
```{r TOOTHGROWTH SUMMARY, echo=T, message=F, warning=F}
summary(ToothGrowth)
```
It is possible to see that that are 3 possible values from **dose** variable: `r unique(ToothGrowth$dose)`. 
```{r BOXPLOT BY DOSIS, echo=T, message=F, warning=F, fig.width=5, fig.asp=.62}
boxplot(len ~ as.factor(dose), data = ToothGrowth, xlab = "dose")
```
There seems to be a considerable difference on variable **len** as **dose** increases. To make sure, it is important to construct the confidence intervals (with 95%) and make sure the extremes does not superimpose themselves.
```{r CONFIDENCE INTERVALS, echo=T, message=F, warning=F}
ToothGrowth %>%
        group_by(dose) %>%
        summarise(LCIL = mean(len) - qt(0.975, df = (length(len)-1) * sd(len)/sqrt((length(len)))),
                  UCIL = mean(len) + qt(0.975, df = (length(len)-1)) * sd(len)/sqrt((length(len))))
```
As no UCIL (Upper confidence interval limit) is bigger than LCIL (Lower confidence interval limit), it is possible to say that all diferente values of **dose** procuce statiscal different means for **len**, considering a alpha of 5%.