---
title: "Notebook for Statistical Inference Course Project 01"
author: "Samuel B Baco"
output:
  pdf_document: default
  html_notebook: default
---
# PACKAGE LOADING
```{r PACKAGE_LOADING, echo=TRUE, message=FALSE, warning=F}
library(tidyverse)
library(ggpubr)
library(datasets)
set.seed(123456)
```
# PART 01: SIMULATION EXERCISE

### 1.1 OVERVIEW

This report is related to course project (week 04) for Johns Hopkins Coursera Statistical Inference classes. The main object is to study the Exponential Distribution e compare it with the Central Limit Theorem.

### 1.2 SIMULATIONS

Below it is possible to find the code and results to simulate 1000 exponentials, all having sample size if 40 and lambda of 0.2. The result was saved at **exponentials** variable.

```{r SIMULATIONS, echo = TRUE}
lambda <- 0.2
n <- 40 # samples for each distribution
N <- 1000 # totals of distribution
exponentials <- replicate(N, rexp(n, lambda))
```

### 1.3 SAMPLE MEAN VS THEORETICAL MEAN

For *lambda* = 0.2, the exponential distributions has an theoretical mean of 5 (*1/lambda*). To calculate the mean for all 1000 generated distributions, the *apply* function will be used.
```{r SAMPLE MEAN, echo = TRUE, message = FALSE}
sMean <- as.data.frame(apply(exponentials, 2, mean))
names(sMean) <- c("mean.exp")
summary(sMean)
```
Comparing the sample mean of theoretical mean, they are pretty close (`r mean(sMean$mean.exp)` vs 5). 

## 1.4 SAMPLE VARIANCE VS THEORETICAL VARIANCE

Using the same dataframe created at **exponentials** and the same method (*apply*), the variances for all 1000 distributions were calculated. The theoretical variance for this example is 25 (*[1/lamda]²*).
```{r SAMPLE VARIANCE, echo=TRUE}
sVar <- as.data.frame(apply(exponentials, 2, var))
names(sVar) <- c("variance.exp")
summary(sVar)
```
Comparing the sample variance with theoretical variance, they are pretty close either (`r mean(sVar$variance.exp)` vs 25).

## 1.5 DISTRIBUTION

In this section, the normality of the data will be investigated. From Central Limit theorem, it is know that a distribution of means is always normal. 

### Sample Mean
```{r SAMPLE MEAN HISTOGRAM, echo = F, message = FALSE, fig.width=4, fig.asp=.62}
g <- ggplot(data = sMean, mapping = aes(x = mean.exp))
g + geom_histogram(aes(y = ..density..)) + 
        labs (x = "Mean",y = " ", title = "Mean histogram for exponential distribution", 
              subtitle = paste("Density for", N, "distributions with lamba =", 1/lambda, "and sample =", n, sep = " ")) +
        geom_vline(aes(xintercept = 1/lambda, color = "darkblue"), size = 2) +
        scale_color_identity(guide = "legend", name = "Theoretical Mean", labels = 1/lambda) +
        stat_function(fun = dnorm, args = list(mean = mean(sMean$mean.exp), sd = sd(sMean$mean.exp)), color = "darkred")

```

Doing a graphical analysis, it seems that the distribution of the sample mean is pratically normal. To make sure, it is necessary to run a normality test.
```{r MEANS NORMALLITY TEST, echo=T, message=F, fig.width=3, fig.asp=.62}
ggqqplot(sMean$mean.exp)
```

It is possible to see the data follows a normal distribution.
\pagebreak

# APPENDIX

## Code for Exponential Means graphic

```{r SAMPLE MEAN HISTOGRAM CODE, eval = F, echo = T, message=F, warning=F}
g <- ggplot(data = sMean, mapping = aes(x = mean.exp))
g + geom_histogram(aes(y = ..density..)) + 
        labs (x = "Mean",y = " ", title = "Mean histogram for exponential distribution", 
              subtitle = paste("Density for", N, "distributions with lamba =", 1/lambda, "and sample =", n, sep = " ")) +
        geom_vline(aes(xintercept = 1/lambda, color = "darkblue"), size = 2) +
        scale_color_identity(guide = "legend", name = "Theoretical Mean", labels = 1/lambda) +
        stat_function(fun = dnorm, args = list(mean = mean(sMean$mean.exp), sd = sd(sMean$mean.exp)), color = "darkred")
print("Don't run me")

```

